Spike spectra for recurrences

### 2022-10-20 Manuscript/inter_spike_manuscript_mdpi.tex
abstract

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{A novel kind of power spectrum is constructed, the \textit{inter-spike spectrum}, which transforms any signal into its spike-frequency domain. This method 
clearly shows the apparent cycles in the data and overcomes the problems for spike-train-like signals when using the obvious idea of Fourier-transforming it. 
We invent this instructive approach with the 
idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter-spike spectrum can be interpreted as a nonlinear power 
spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is robust to noise and is able to detect and analyze bifurcations.}

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{In reccurance analysis, the $\tau$-recurrence rate encodes crucial information on the cycles of the underlying high-dimensional timeseries. It plays a similar role to the autocorrelation for scalar time-series in encoding temporal correlations, however, it is not amenable to analysis by Fourier decomposition. Thus there is no natural analogue to the power spectrum in recurrence analysis. In this paper we introduce a novel way to decompose the $\tau$-recurrence rate using an overcomplete basis of Dirac combs together with a sparsity regularization. We show that this decomposition, the \textit{inter-spike spectrum}, naturally provides an analogue to the power spectrum for recurrence analysis. We show that the inter-spike spectrum correctly identifies patterns and transitions in the underlying timeseries in a wide variety of examples and is robust to measurement noise.

A novel kind of power spectrum is constructed, the \textit{inter-spike spectrum}, which transforms any signal into its spike-frequency domain. This method 
clearly shows the apparent cycles in the data and overcomes the problems for spike-train-like signals when using the obvious idea of Fourier-transforming it. 
We invent this instructive approach with the 
idea of transforming the $\tau$-recurrence rate of a recurrence plot (RP), which often has a spiky appearance. The $\tau$-recurrence rate is the density of recurrence points 
along diagonals of the RP, which are parallel to the main diagonal with a distance of $\tau$. In this context the inter-spike spectrum can be interpreted as a nonlinear power 
spectrum of a potentially high dimensional system which constitutes the RP. The proposed measure is robust to noise and is able to detect and analyze bifurcations.}
